---
title: "Rmd/Rnw rename code chunks vignette"
#subtitle: "(Re)Name code chunks with specified chunk name, ignoring differently-named chunks"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 5
    number_sections: true
vignette: >
  %\VignetteIndexEntry{Rmd/Rnw rename code chunks vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
#always_allow_html: yes
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(erikmisc)
```

Read an Rmd file, update existing prefixed code chunks, and renumber.
```{r}
e_rmd_name_chunks(
    fn_in             = "test_in.Rmd"
  , fn_out            = "test_out.Rmd"
  , prefix_chunk_name = "chunk-"
  )
```

Review the input and output files to see how the chunk header names have been
updated.

```
test_in.Rmd
 8 : ```{r setup, include=FALSE}
22 : ```{R cars}
28 : ```{r}
32 : ```{  R  }
36 : ```{r, echo=FALSE}
40 : ```{   R, eval=FALSE}
44 : ```{r   , eval=FALSE}
48 : ```{r chunk-2, eval=FALSE}
52 : ```{r chunk-XXX  , eval=FALSE}
56 : ```{R chunk-XXX2 , eval=FALSE}
60 : ```{R chunk-XXX3 , eval=FALSE}
64 : ```{r chunk-XXX4 , eval=FALSE}
68 : ```{r chunk-XXX5 , eval=FALSE}
72 : ```{r chunk-XXX6 , eval=FALSE}
81 : ```{r pressure, echo=FALSE}

# These code chunk headers were changed to those below:

test_out.Rmd
 8 : ```{r setup, include=FALSE}
22 : ```{r cars}
28 : ```{r chunk-01}
32 : ```{r chunk-02}
36 : ```{r chunk-03, echo=FALSE}
40 : ```{r chunk-04, eval=FALSE}
44 : ```{r chunk-05, eval=FALSE}
48 : ```{r chunk-06, eval=FALSE}
52 : ```{r chunk-07, eval=FALSE}
56 : ```{r chunk-08, eval=FALSE}
60 : ```{r chunk-09, eval=FALSE}
64 : ```{r chunk-10, eval=FALSE}
68 : ```{r chunk-11, eval=FALSE}
72 : ```{r chunk-12, eval=FALSE}
81 : ```{r pressure, echo=FALSE}
```

