---
title: "Plot model contrasts"
subtitle: "Plotting all linear model effects and contrasts"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 5
    number_sections: true
vignette: >
  %\VignetteIndexEntry{Plot model contrasts}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
knitr::opts_chunk$set(fig.height = 6, fig.width = 6)

library(erikmisc)
```


# Plotting model contrasts

The `e_plot_model_contrasts()` function is a wrapper for `emmeans::emmeans()`.
It automates plotting all of the model effects and annotating the plots with estimates and contrasts.
Now it is (almost) effortless to interpret a linear model output.

This has been tested for `lm()` (fixed-effects model) and `lme4::lmer` (mixed fixed- and random-effects model).

## Data

The data variables are labelled and those will be used for plots.

```{r}
str(dat_mtcars_e)
```

## Fit model

This model is specifically chosen to have numeric and categorical main effects,
  and each type of two-way interaction (cat:cat, cat:num, num:num),
  as well as some categories that were not estimable.
"Statistical significance" and "model fit" are not concerns in this demonstration.

```{r}
 # Set specific model with some interactions
form_model <-
  mpg ~ cyl + disp + hp + wt + vs + am + cyl:vs + disp:hp + hp:vs

fit <-
  lm(
    formula = form_model
  , data    = dat_mtcars_e
  )

anova(fit)
```

```{r}
summary(fit)
```

## Plotting model effects and contrasts

### Defaults

See the help `?e_plot_model_contrasts` for the method defaults.
I'm choosing to not print the output (default) while the function runs so that
we can examine the separate lists in the output.

```{r}
fit_contrasts <-
  e_plot_model_contrasts(
    fit                = fit
  , dat_cont           = dat_mtcars_e
  , sw_print           = FALSE
  )
```

The output contains lists for tables, plots, and text.

```{r}
fit_contrasts %>% names()
```

#### tables

The `tables` are give the estimates (`$est`) and contrasts (`$cont`) for each
effect in the model not involved with an interaction (argument
`sw_marginal_even_if_interaction` can include all main effects).

```{r}
fit_contrasts$tables  # to print tables
```

#### text

The `text` of the tables are written in a form that can be copy/pasted into a
report and worked into sentences.

```{r}
fit_contrasts$text    # to print caption text
```

#### plots

The `plots` illustrate the effects with labelled plots and the `text` in the
caption for a single complete result.

From the help:

    Plot interpretation:
    This EMM plot (Estimated Marginal Means, aka Least-Squares Means) is only
    available when conditioning on one variable. The blue bars are confidence
    intervals for the EMMs; don't ever use confidence intervals for EMMs to perform
    comparisons --- they can be very misleading. The red arrows are for the
    comparisons between means; the degree to which the "comparison arrows" overlap
    reflects as much as possible the significance of the comparison of the two
    estimates. If an arrow from one mean overlaps an arrow from another group, the
    difference is not significant, based on the adjust setting (which defaults to
    "tukey").

For two-way interactions, both directions of conditioning are plotted: `var1|var2` and `var2|var1`.
Furthermore, each plot is made separately and given in a side-by-side or above-below orientation.
Use `ggsave()` for a specific plot object and choose a good `width` and `height` to make it look beautiful.

```{r}
fit_contrasts$plots   # to print plots
```

### Some options

Below I use some options to specify one two-way interaction, stack them above-below ("tall"), and exclude the table in the plot.

```{r, fig.height = 7, fig.width = 6}
fit_contrasts <-
  e_plot_model_contrasts(
    fit                = fit
  , dat_cont           = dat_mtcars_e
  , choose_contrasts   = "cyl:vs"
  , sw_print           = FALSE
  , sw_table_in_plot   = FALSE
  , sw_TWI_plots_keep  = c("singles", "both", "all")[2]
  , sw_TWI_both_orientation = c("wide", "tall")[2]
  )

fit_contrasts$plots   # to print plots
fit_contrasts$tables  # to print tables
fit_contrasts$text    # to print caption text
```

