---
title: "Best subset selection"
subtitle: "Plotting all linear model effects and contrasts"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 5
    number_sections: true
vignette: >
  %\VignetteIndexEntry{Best subset selection}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
#always_allow_html: yes
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(erikmisc)
```


# Best subset selection

The `e_lm_bestsubset()` function is a wrapper for `leaps::regsubsets()`.
It takes a "full model" formula and returns the best `nbest` models of each model size (1 predictor, 2 predictors, etc.),
in a table that's easy to review.

NOTE: This does not assess model fit!  For that, see `e_lm_diag_plots()`.

## Data

Fit the best 3 models of each model size.
For each column, a 1 or 0 indicates the variable is or is not in the model.
The model `SIZE` is how many terms are in the model (not counting the intercept).
Other model statistics are provided and the results are sorted by BIC (lowest BIC is preferred, best on top).

```{r}
tab_best <-
  e_lm_bestsubset(
    form  = formula(mpg ~ cyl + disp + hp + gear)
  , dat   = datasets::mtcars
  , nbest = 3
  )

  ### consider these options if you temporarily need a wider output in your Rmd output
  # op <- options(); # saving old options
  # options(width=100) # setting command window output text width wider
tab_best %>% print(n = Inf, width = Inf)
  # options(op); # reset (all) initial options
```

```{r}
# print into RStudio viewer or a pretty table in Rmd output
# include "always_allow_html: yes" in Rmd yaml header
tab_best %>% e_table_print(sw_kable_format  = c("simple", "kbl", "html", "latex", "doc")[1])
```
