---
title: "Multiple regression using erikmisc"
subtitle: "Plotting all linear model effects and contrasts"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 5
    number_sections: true
vignette: >
  %\VignetteIndexEntry{Multiple regression using erikmisc}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

* This is an example of multiple regression using the functions in the `erikmisc` package.

```{r setup}
knitr::opts_chunk$set(fig.width = 6, fig.height = 6)

library(erikmisc)
```

# Load data

* Calculating `mpg_log2` and dropping two extremely influential observations.
* Creating some missing values to illustrate missing values plot.

```{r}
data(dat_mtcars_e)
dat_mtcars_e <-
  dat_mtcars_e %>%
  dplyr::mutate(
    mpg_log2 = mpg %>% log2()
  ) %>%
  dplyr::slice(
    -c(29, 31)
  ) %>%
  dplyr::select(
    model, mpg, mpg_log2
  , tidyselect::everything()
  )

# create some missing values in variable we won't use in the regression model
var_missing <- c("qsec", "carb", "am")
prop_missing = 0.25
n_missing = sample.int(n = nrow(dat_mtcars_e) * length(var_missing), size = round( prop_missing * nrow(dat_mtcars_e) * length(var_missing)))
ind_missing = expand.grid(1:dim(dat_mtcars_e)[1], 1:length(var_missing))[n_missing, ]
for (i_row in seq_len(nrow(ind_missing))) {
  dat_mtcars_e[ind_missing[i_row,1], var_missing[ind_missing[i_row,2]] ] <- NA
}

str(dat_mtcars_e)
#summary(dat_mtcars_e)
```

## Plot missing values

```{r}
e_plot_missing(
    dat_plot       = dat_mtcars_e
  , var_group      = "cyl"
  , sw_group_sort  = TRUE
  , var2_sort      = "disp"
  )
```

# Fit full model

* This model is chosen to have numeric and categorical main effects,
  and at least two of possible types of two-way interaction (cat:cat, cat:num, num:num).

```{r}
 # Set model with all two-way interactions
#form_model <-
#  mpg ~ (cyl + disp + hp + wt + vs + am + gear)^2
#form_model <-
#  mpg ~ (cyl + disp + hp + wt + am)^2
form_model <-
  mpg_log2 ~ (cyl + disp + hp + wt)^2

# fit model
lm_fit <-
  lm(
    formula = form_model
  , data    = dat_mtcars_e
  )
```

* Anova shows of the many two-way interactions in the model, almost none are significant.
* We can greatly simplify this model.

```{r}
car::Anova(lm_fit, type = 3)
```

* Summary shows that all effects were estimable.

```{r}
summary(lm_fit)
```

## Assess model assumptions

* On full dataset:
  * QQ-plot shows roughly normal (outliers in the direction of high kurtosis, "leptokurtosis").
  * Cook's distance shows two highly influential observations.
  * Cook's vs leverage plot shows that these influential observations (y-axis) are because of their high leverage (outlier in the x-direction).
  * Box-Cox suggests a lambda = 0, log(y) transformation will improve normality.
* After log2 transform and dropping two extremely influential observations (below)...
  * Cook's distance shows two or three influential observations.

_Recall, our goal is not to resolve all issues, but to illustrate the functions in the `erikmisc` package._

```{r, fig.width = 8, fig.height = 3}
e_plot_lm_diagostics(
    fit = lm_fit
  , sw_plot_set = "simple"
  )
```

# Best subset selection

* The best model by BIC has `hp` and `wt`.

```{r}
tab_lm_best <-
  e_lm_bestsubset(
    form  = form_model
  , dat   = dat_mtcars_e
  , nbest = 4
  )

  ### consider these options if you temporarily need a wider output in your Rmd output
  op <- options(); # saving old options
  options(width = 200) # setting command window output text width wider
tab_lm_best %>% print(n = Inf, width = Inf)
  options(op); # reset (all) initial options
```

# Fit reduced model

* For demonstration purposes, I'm choosing Model 9 because it has both `factor:numeric` and `numeric:numeric` interactions.

```{r}
tab_lm_best[9, ] %>% print(width = Inf)

form_model_red <-
  mpg_log2 ~ cyl + hp + wt + cyl:wt + hp:wt

# fit reduced model
lm_fit_red <-
  lm(
    formula = form_model_red
  , data    = dat_mtcars_e
  )
car::Anova(lm_fit_red, type = 3)
summary(lm_fit_red)
```

## Assess model assumptions

* Cook's distance shows Observation 17 is influential.

```{r, fig.width = 8, fig.height = 3}
e_plot_lm_diagostics(
    fit = lm_fit_red
  , sw_plot_set = "simpleAV"
  )
```

# Plot model contrasts

* See the help `?e_plot_model_contrasts` for the method defaults.
* I'm choosing to not print the output (default) while the function runs so that
  we can examine the separate lists in the output.

```{r, fig.width = 10, fig.height = 8}
lm_fit_contrasts <-
  e_plot_model_contrasts(
    fit                 = lm_fit_red
  , dat_cont            = dat_mtcars_e
  , adjust_method       = c("none", "tukey", "scheffe", "sidak", "bonferroni", "dunnettx", "mvt")[2]
  , sw_print            = FALSE
  , sw_TWI_plots_keep   = c("singles", "both", "all")[2]
  )
```

* The output contains lists for tables, plots, and text.

```{r}
lm_fit_contrasts %>% names()
```

## tables

* The `tables` are give the estimates (`$est`) and contrasts (`$cont`) for each
  effect in the model not involved with an interaction (argument
  `sw_marginal_even_if_interaction` can include all main effects).

```{r}
# to print tables
lm_fit_contrasts$tables
```

## text

* The `text` of the tables are written in a form that can be copy/pasted into a
  report and worked into sentences.

```{r}
# to print caption text
#lm_fit_contrasts$text
```

## plots

* The `plots` illustrate the effects with labelled plots and the `text` in the
  caption for a single complete result.

```{r, fig.width = 8, fig.height = 4}
# These are the two plots
lm_fit_contrasts$plots$`cyl:wt`$both
lm_fit_contrasts$plots$`hp:wt`$both
```

