---
title: "All model subsets"
subtitle: "Best subset selection for all binary subsets of selected variables"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 5
    number_sections: true
vignette: >
  %\VignetteIndexEntry{All model subsets}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This has a few examples of specifying model subsets and fitting linear models
in order to choose the "best" model via a model selection criterion.

```{r setup}
library(erikmisc)
library(tidyverse)
```

# Using `e_model_all_subsets_formula`

In this example, variables "a" and "b" are always in the model,
  while variables "c" and "d" are subject to model selection up to a second-order model.

* Maximal model: `z ~ a + b + c + d + c^2 + d^2 + c:d`
* Minimal model: `z ~ a + b`

In biostatistical or epidemiological applications,
  certain demographic variables such as age, sex, and race
  are often included regardless of "statistical signicance",
  and other variables are subject to selection.

```{r}
# Two variables always in the model with a second-order model of two others
e_model_all_subsets_formula(
    var_formula = NULL
  , x_var_names = list(c("a", "b"), c("c", "d"))
  , y_var_name  = "z"
  , max_scope   = list("always", "SO")
  , sw_return   = c("formula", "table")[2]
  )
```

# Applications

# Linear model fit and model selection

_Note that in these examples we are ignoring model fit;
  see `e_plot_lm_diagostics` for diagnostics._

We'll fit all the linear models between the maximal and minimal models:

* Maximal model: `mpg ~ cyl + carb + disp + hp + disp:hp + wt + vs + am + gear`
* Minimal model: `mpg ~ cyl + carb`

Then we select the model with the lowest BIC statistic.

```{r}
dat_mtcars_e %>% head(3)

# table of the candidate models
form_table <-
  e_model_all_subsets_formula(
      var_formula = NULL
    , x_var_names = list(c("cyl", "carb"), c("disp", "hp"), c("wt", "vs", "am", "gear"))
    , y_var_name  = "mpg"
    , max_scope   = list("always", "TWI", "FO")
    , sw_return   = c("formula", "table")[2]
    )
print(form_table)

# generate candidate model formulas
form_list <-
  e_model_all_subsets_formula(
      var_formula = NULL
    , x_var_names = list(c("cyl", "carb"), c("disp", "hp"), c("wt", "vs", "am", "gear"))
    , y_var_name  = "mpg"
    , max_scope   = list("always", "TWI", "FO")
    , sw_return   = c("formula", "table")[1]
    )

# fit models and calculate model selection criteria
lm_fit_list <- list()
lm_fit_list_criteria <- list()

for (i_form in seq_len(length(form_list))) {
  # model fit
  lm_fit_list[[ i_form ]] <-
    stats::lm(
      formula = form_list[[ i_form ]]
    , data    = dat_mtcars_e
    )

  # model selection criteria
  lm_fit_list_criteria[[ i_form ]] <-
    e_lm_model_criteria(
      lm_fit    = lm_fit_list[[ i_form ]]
    , dat_fit   = dat_mtcars_e
    , model_id  = i_form
    )
}

# sort the models by a criterion
lm_fit_criteria <-
  lm_fit_list_criteria %>%
  dplyr::bind_rows() %>%
  dplyr::arrange(
    # arrange models by this model criterion
    bic               # ascending by bic (best on top)
    # dplyr::desc(r2)   # descending by r2 (best on top)
  )

# print the head and tail of fit criteria table
lm_fit_criteria %>% head()
lm_fit_criteria %>% tail(3)

# best model ID
lm_fit_criteria$model_id[1]

# summarize "best" model
lm_fit_list[[ lm_fit_criteria$model_id[1] ]] %>%
  car::Anova(type = 3)
lm_fit_list[[ lm_fit_criteria$model_id[1] ]] %>%
  summary()
```



# Longitudinal linear model fit and model selection

For the purpose of illustrating the method, variables `x1`, `x2`, and `x3` are
  generated as indepedent standard normal random variables.
The correct model would remove all of those variables and keep only the
  fixed effect of `Type` and the random effect `(1 | Subject)`.

We'll fit all the longitudinal linear models between the maximal and minimal models:

* Maximal model: `Effort ~ Type + (1 | Subject) + x1 + x2 + x3 + x1^2 + x2^2 + x3^3 + x1:x2 + x1:x3 + x2:x3`
* Minimal model: `Effort ~ Type + (1 | Subject)`

Then we select the model with the lowest cAIC statistic.

```{r}
set.seed(76543)

dat_ergoStool_e <-
  dat_ergoStool_e %>%
  # add a few noise variables just to use for model selection example
  dplyr::mutate(
    x1 = stats::rnorm(n = dplyr::n())
  , x2 = stats::rnorm(n = dplyr::n())
  , x3 = stats::rnorm(n = dplyr::n())
  )
dat_ergoStool_e %>% head(8)

# table of the candidate models
form_table <-
  e_model_all_subsets_formula(
      var_formula = NULL
    , x_var_names = list(c("Type", "(1 | Subject)"), c("x1", "x2", "x3"))
    , y_var_name  = "Effort"
    , max_scope   = list("always", "SO")
    , sw_return   = c("formula", "table")[2]
    )
print(form_table)

# generate candidate model formulas
form_list <-
  e_model_all_subsets_formula(
      var_formula = NULL
    , x_var_names = list(c("Type", "(1 | Subject)"), c("x1", "x2", "x3"))
    , y_var_name  = "Effort"
    , max_scope   = list("always", "SO")
    , sw_return   = c("formula", "table")[1]
    )

# fit models and calculate model selection criteria
lmer_fit_list <- list()
lmer_fit_list_criteria <- list()

for (i_form in seq_len(length(form_list))) {
  # model fit
  lmer_fit_list[[ i_form ]] <-
    lme4::lmer(
      formula = form_list[[ i_form ]]
    , REML    = TRUE
    , data    = dat_ergoStool_e
    )

  # model selection criteria
  lmer_fit_list_criteria[[ i_form ]] <-
    e_lm_model_criteria(
      lm_fit    = lmer_fit_list[[ i_form ]]
    , dat_fit   = dat_ergoStool_e
    , model_id  = i_form
    )
}

# sort the models by a criterion
lmer_fit_criteria <-
  lmer_fit_list_criteria %>%
  dplyr::bind_rows() %>%
  dplyr::arrange(
    # arrange models by this model criterion
    caic              # ascending by caic (best on top)
    # dplyr::desc(r2)   # descending by r2 (best on top)
  )

# print the head and tail of fit criteria table
lmer_fit_criteria %>% head()
lmer_fit_criteria %>% tail(3)

# best model ID
lmer_fit_criteria$model_id[1]

# summarize "best" model
lmer_fit_list[[ lmer_fit_criteria$model_id[1] ]] %>%
  car::Anova(type = 3)
lmer_fit_list[[ lmer_fit_criteria$model_id[1] ]] %>%
  summary()
```


# Longitudinal linear (in parallel, for 200+ models)

```{r, eval=FALSE}
## Set up parallel backend
  ## From: https://www.blasbenito.com/post/02_parallelizing_loops_with_r/
  library(doParallel)
  # use n-1 cores
  n_cores <- parallel::detectCores() - 1
  # create the cluster
  my_cluster <-
    parallel::makeCluster(
      spec = n_cores
    , type = "PSOCK"
    )
  # check cluster definition (optional)
  print(my_cluster)
  # register it to be used by %dopar%
  doParallel::registerDoParallel(cl = my_cluster)
  # check if it is registered (optional)
  foreach::getDoParRegistered()
  # how many workers are available? (optional)
  foreach::getDoParWorkers()


# table of the candidate models
form_table <-
  e_model_all_subsets_formula(
      var_formula = NULL
    , x_var_names = list(c("Type", "(1 | Subject)"), c("x1", "x2", "x3"))
    , y_var_name  = "Effort"
    , max_scope   = list("always", "SO")
    , sw_return   = c("formula", "table")[2]
    )
print(form_table)

# generate candidate model formulas
form_list <-
  e_model_all_subsets_formula(
      var_formula = NULL
    , x_var_names = list(c("Type", "(1 | Subject)"), c("x1", "x2", "x3"))
    , y_var_name  = "Effort"
    , max_scope   = list("always", "SO")
    , sw_return   = c("formula", "table")[1]
    )

# fit models and calculate model selection criteria
lmer_fit_list <-
  foreach::foreach(
    i_form = seq_len(length(form_list))
  , .combine = "c"
  ) %dopar% {
    # model fit
    lme4::lmer(
      formula = form_list[[ i_form ]]
    , REML    = TRUE
    , data    = dat_ergoStool_e
    )
  }

lmer_fit_list_criteria <-
  foreach::foreach(
    i_form = seq_len(length(form_list))
  #, .combine = "c"
  ) %dopar% {
    # model selection criteria
    e_lm_model_criteria(
      lm_fit    = lmer_fit_list[[ i_form ]]
    , dat_fit   = dat_ergoStool_e
    , model_id  = i_form
    )
  }



# sort the models by a criterion
lmer_fit_criteria <-
  lmer_fit_list_criteria %>%
  dplyr::bind_rows() %>%
  dplyr::arrange(
    # arrange models by this model criterion
    caic              # ascending by caic (best on top)
    # dplyr::desc(r2)   # descending by r2 (best on top)
  )

# print the head and tail of fit criteria table
lmer_fit_criteria %>% head()
lmer_fit_criteria %>% tail(3)

# best model ID
lmer_fit_criteria$model_id[1]

# summarize "best" model
lmer_fit_list[[ lmer_fit_criteria$model_id[1] ]] %>%
  car::Anova(type = 3)
lmer_fit_list[[ lmer_fit_criteria$model_id[1] ]] %>%
  summary()


## stop cluster
parallel::stopCluster(cl = my_cluster)
```

